Exploring Determinants of Child Mortality Globally (2022): The Role of Vaccination, Healthcare Spending, and Nutrition

### 2025-04-15 paper/paper.qmd
Result

Here we will investigate two loss functions, ranknet loss and focal loss. Ranknet loss is sepcified as follow: Given training data $x_1$, $x_2$, .... , $x_n$, and their rankings, $y_1$, $y_2$, ..., $y_n$, we will investigate all combinations of ordered pairs ($x_i$, $x_j$). The total loss is: 
$$
L = \sum_{i < j} - y_{ij} \log P(i > j) - (1 - y_{ij}) \log (1 - P(i > j))
$$
Where $y_ij$ is 1 if $x_i$ is ranked higher than $y_j$, 0 if $x_i$ is ranked lower than $y_j$ (here we will assume no $x_i$ can have the same ranking as $x_j$ if i is not equal to j). $P(i > j) = \frac{1}{1 + e^{-(s(x_i) - s(x_j))}}$ is the probability that $x_i$ is ranked higher than $x_j$. s() is the rankning score assigned to i, higher ranking cores means higher ranking. This loss aims to peanalize the pairs of data $x_i$ and $x_j$ where $x_i$ should be ranked higher than $x_j$ but instead $x_i$ is ranked lower by the model for $x_j$.

@@

>>>>

Here we will investigate two loss functions, ranknet loss and focal loss. Ranknet loss is sepcified as follow: Given training data $x_1$, $x_2$, .... , $x_n$, and their rankings, $y_1$, $y_2$, ..., $y_n$, we will investigate all combinations of ordered pairs ($x_i$, $x_j$) where their labels $y_i$ and $y_j$ are different. The total loss is: 
$$
L = \sum_{i < j} - y_{ij} \log P(i > j) - (1 - y_{ij}) \log (1 - P(i > j))
$$
Where $y_ij$ is 1 if $x_i$ is ranked higher than $y_j$, 0 if $x_i$ is ranked lower than $y_j$ (here we will assume no $x_i$ can have the same ranking as $x_j$ if i is not equal to j). $P(i > j) = \frac{1}{1 + e^{-(s(x_i) - s(x_j))}}$ is the probability that $x_i$ is ranked higher than $x_j$. s() is the rankning score assigned to i, higher ranking cores means higher ranking. This loss aims to peanalize the pairs of data $x_i$ and $x_j$ where $x_i$ should be ranked higher than $x_j$ but instead $x_i$ is ranked lower by the model for $x_j$.

@@

To test the two loss functions useing the new metric we build a one layer neuronetwork with 40 input neurons, 128 hidden neurons and one output which is the score of the input data in pytorch. We split the dataset into training and testing parts with the ratio of 7:3. To gain efficiency in training, we split the training data into minibatches of size 128 and after finetuning set the training epohs be 100 to avoid overtraining. We also used a neuronet work with crossentropy loss as a conterpart. We ran 10 independent groups and the result is the following:
